\title{SVD-based sound speed field estimation with null-space exploiting regularization
	\thanks{This work was supported by Petrobras, grant number XXXXX.}
	}
%\title{SVD}
%
%\title{Directional beamforming with direction of arrival steering and noise covariance matrix estimation}

\author{\IEEEauthorblockN{1\textsuperscript{st} Vitor G. P. Curtarelli}
	\IEEEauthorblockA{%
		\textit{Dept. of Electrical}\\\textit{and Electronic Engineering}\\\textit{UFSC}\\
		Florianópolis, Brazil \\
		\url{vitor.curtarelli@posgrad.ufsc.br}}
	\and
	\IEEEauthorblockN{2\textsuperscript{nd} Anderson W. Spengler}
	\IEEEauthorblockA{%
		\textit{Dept. of Electrical}\\\textit{and Electronic Engineering}\\\textit{UFSC} \\
		Florianópolis, Brazil \\
		\url{anderson.spengler@ufsc.br}}
	\and
	\IEEEauthorblockN{3\textsuperscript{rd} Stephan Paul}
	\IEEEauthorblockA{%
		\textit{Dept. of Mechanical Engineering} \\
		\textit{UFSC}\\
		Florianópolis, Brazil \\
		\url{stephan.paul@ufsc.br}}
}

\maketitle

\begin{abstract}
    This paper proposes a singular-value decomposition -based approach for inverse mapping in sound speed field reconstruction in multi-dimensional environments, using travel-time information. This method exploits the null space of the forward operator to perform regularization, replacing the traditional additive regularization term. It enables minimization of the primary cost function while controlling solution properties through a secondary optimization. For comparisons, a conventional (literature-based) approach for travel-time tomography is also presented. Simulations in ocean-like 2D environments show comparable performance between methods under standard device configurations, while the proposed approach performs better in more practical configurations, although with overall limited accuracy.
\end{abstract}

\begin{IEEEkeywords}
	Travel-time tomography, singular value decomposition, sound speed field estimation
\end{IEEEkeywords}