\section{Introduction}

Although the speed of sound $c$ in water is considered constant at around $1500 \si{\meter}/\si{\second}$ \cite{delgrosso_new_1974}, this approximation is invalid for large-scale environments \cite{pinson_sound_2010}. Many properties can affect the speed of sound in water, such as pressure, temperature, salinity (abbreviated TPS), and water currents \cite{medwin_fundamentals_1999,mcdougall_getting_2011,delgrosso_new_1974}, and as these properties vary over the body of water, so does the velocity, creating a sound speed profile (SSP) in a 1D scenario \cite{li_acoustic_2019,pinson_sound_2010}, or a sound speed field (SSF)\footnote{Although the acronym SSF means ``sound speed field'', it is more practical to work with the slowness, the inverse of speed. The ``sound slowness field'' will also be referenced by the acronym SSF, as they are inversely proportional.} in a multi-dimensional case.

The knowledge of the SSF enables estimating the wavefunction and its behavior as it travels through this medium \cite{li_acoustic_2019,pinson_sound_2010}. By Snell's law and Fermat's principle of least time \cite{perrine_fermats_1961}, a change in speed of a wave along its trajectory also changes its direction. This causes curves, refractions, reflections, and formation/dispersion of beams \cite{cornuelle_ocean_2008}. For a sufficiently large body of water, these wavepath transformations are noticeable, and need to be accounted for. By treating an omnidirectional wave as a cluster of unidirectional waves, we can calculate the travel-time between the source and any point within the medium, as well as its wavefront's profile. The knowledge of the path taken by an ensemble of rays is of great importance in the fields of seismology \cite{bormann_seismic_2012} and oceanography \cite{cornuelle_ocean_2008}, and the estimation of the SSF within the environment is relevant for acoustics \cite{holstein_acoustic_2004}, medicine \cite{hormati_robust_2010}, and others.

Initial works in the topic of estimating the velocity field in water aimed at estimating the TPS properties, as well as the relationship between them and the sound speed in water. Works such as \cite{medwin_fundamentals_1999,mcdougall_getting_2011,delgrosso_new_1974} operated under this framework, with some still being used today to model the behavior of water currents on a global scale.

While traditional techniques used the environmental TPS properties to estimate the SSF, most modern SSF estimation approaches avoid estimating these properties, and treat this SSF estimation as an inversion problem. That is, the knowledge of some information regarding the wave paths within the environment permits the estimation of the SSF in the medium. Different techniques approach this inversion problem by measuring different wave properties, and taking distinct mathematical models. Such techniques encompass Travel-Time Tomography \cite{stefanov_travel_2019}, Matched Field Processing \cite{gemba_adaptive_2017} and Matched Field Inversion \cite{dosso_bayesian_2011}, Full Waveform Inversion \cite{virieux_overview_2009}, a mix among these \cite{treister_full_2017}, or a fusion with machine learning \cite{zhang_bayesian_2021}.

Among these, the Travel-Time Tomography (TTT) is known for its simplicity and versatility \cite{stefanov_travel_2019,aki_determination_1977,phillips_traveltime_1991,tang_travel_2024}. It primarily uses the difference in arrival time of the acoustic signals at the receivers, requiring a spatial distribution of sources and receivers for appropriate spatial resolution and sufficient spatial information. In general, the SSF is discretized in cells (or layers in the 1D scenario), wherein the sound speed is considered constant. The SSF's inversion iterates alternately between calculating the paths and lengths in each cell via solving the Eikonal equation \cite{au_geometrical_2005} (generally through the Fast Marching Method \cite{tang_travel_2024}) (called the forward map); and estimating the speed in each cell (called the inverse map), which is tackled by the TTT. As previously cited, the TTT is known for its simplicity, versatility, and low computational complexity. However, it depends on a precise positioning and syncing of the devices.

Many modifications on the TTT have been proposed, aiming at improving the quality of the inverse map estimation. Additions such as regularization \cite{aki_determination_1977}, blurring \cite{ali_opensource_2019}, and conjugate gradient minimization \cite{tang_travel_2024} (among others \cite{jeong_investigating_2024,hormati_robust_2010}) have been proposed and studied. Each of these aim to adapt the inverse problem by correcting different issues that may appear, such as ill-posedness, physics-defiant models, or incoherent solutions.

In this paper, we propose a new approach, employing the Singular-Value Decomposition (SVD) and the Moore-Penrose pseudo-inverse (from now on called p-inverse) for the primary estimation problem, as well as a null-space exploiting step for achieving a minimum on a secondary (regularizing) objective function, with minimal impact on the SSF estimation. A small discussion on the theoretical differences between the proposed SVD method and the traditional TTT was also presented, highlighting computational complexity differences, and scenarios where the proposed method can be computationally optimized.

Through simulations emulating real scenarios, results show that both compared frameworks --- through the traditional TTT and the proposed SVD approach --- fare similarly in a scenario closer to the one found in the literature (regarding the devices' positioning). Meanwhile, in a device distribution more aligned with the practical constraints, the SVD method had a noticeably better performance in terms of SSF estimation, although it still underperformed when compared to the literature-based scenario.

%This paper is organized as follows: in \cref{sec:ssf_estimation} we present the inverse problem's mathematical model as well as its physical underlying interpretation, and establish the standard literature technique to which ours will be compared. In \cref{sec:svd_pseudo_ttt} we present the techniques to be used and develop the new proposed model, bringing a discussion on its theoretical advantages and disadvantages. In \cref{sec:simulation_and_results} we present the simulation conditions for comparing the presented techniques, along with the results obtained through these simulations. Finally, \cref{sec:conclusion} concludes the paper.

% Although the acronym SSF means ``sound speed field'', mathematically it is more practical to model the problem via the slowness, the inverse of speed. Since both the sound speed field and sound slowness field would result in the same acronym, and both are simply inversely proportional, the SSF acronym will be used indistinguishably for both. % Throughout the mathematical developments, non-bold variables are scalars, bold lowercase variables are vectors, and bold uppercase variables are matrices.
