\section{SSF estimation and Travel-Time Tomography}
\label{sec:ssf_estimation}
As previously explained, the TTT uses the delay --- travel-time, or first-arrival time --- between the wave's generation at a source and its arrival at a receptor to estimate the path traveled, and therefore the wave's speed along the route taken.

Given a ray that follows a path $\bvr[v]$, and an SSF $c(\bvp)$ as a function of space ($\bvp=[x,y,z]$ in Cartesian coordinates); then the travel time along the curve $\bvr[v]$ is
\begin{equation}
	t_{\bvr[v]} = \int_{\bvr[v]} \frac{1}{c(\bvp)} \dd \bvp.
    \label{eq:sec2:continuous_travel-time}
\end{equation}

This is known as the forward map: given an SSF $c(\bvp)$ and an origin (source) and a destination (receiver), estimate the path $\bvr[v]$ taken, and its travel-time (following Fermat's principle). The inverse problem's objective is to obtain the SSF, given a set of travel-times and paths between different sources and receivers within the environment.

\subsection{Environment modeling and discretization}

Under the TTT technique, the traditional modeling involves discretizing the environment in cells, assuming each cell has a constant speed, and estimating this discrete SSF using the measured travel-times. We divide the environment in a total of $N$ cells of any dimension. Without loss of generality, we will use a 2D environment divided into a $\sz{N_{\z}}{N_{\x}}$ grid. We also assume the presence of $M_{\source}$ sources and $M_{\receiver}$ receivers, resulting in a total of $M = M_{\source} \cdot M_{\receiver}$ rays being produced and traveling through the environment.

Considering this discretization, \cref{eq:sec2:continuous_travel-time} can be written as
\begin{equation}
	t_{m} \approx \sum_{n} r_{m;n} s_n,
	\label{eq:sec2:tm_sum-form}
\end{equation}
where $r_{m;n}$ is the length of the $m$-th ray within the $n$-th cell ($0$ if the given ray doesn't pass through the $n$-th cell), $s_n$ is the $n$-th cell's slowness, and $t_m$ is the $m$-th ray's travel-time. We rewrite \cref{eq:sec2:tm_sum-form} in vector form, as
\begin{equation}
	t_m(\bvs) = \bvr[T]{m}(\bvs) \bvs,
\end{equation}
in which $\bvr{m}(\bvs) \in \re^{\sz{N}{1}}$ corresponds to the lengths in each cell of the $m$-th ray, $\bvs \in \re^{\sz{N}{1}}$ represents the modeled slowness in each cell, and $\tr{\pts{\cdot}}$ stands for the transpose operator. We explicit the dependency of $\bvr{m}(\bvs)$ on $\bvs$, since the ray-bending effects on the $m$-th ray's path will depend on the slowness field, following Fermat's principle. We also highlight that $\bvr{m}(\bvs)$ is a sparse vector, as (for a well-behaved environment) a given ray will pass through only a few cells.

We now let $\bvt(\bvs) \in \re^{\sz{M}{1}}$ as a travel-times vector, for all $M$ considered rays; and $\bvR(\bvs) \in \re^{\sz{M}{N}}$ is a vertical stacking of all $\bvr[T]{m}(\bvs)$. With this, we achieve
\begin{equation}
	\bvt(\bvs) = \bvR(\bvs) \bvs.
	\label{eq:sec2:bvt_product-form}
\end{equation}

We will impose the requirement that $N > M$, ensuring a sufficient resolution on the environment (large $N$) with a small number of measurements (small $M$). We denote $R = \rank{\bvR{i}}$ as the rank of $\bvR{i}$, and it follows that $R \leq M$.

\subsection{SSF estimation}
\label{subsec:sec2:ssf_estimation}

Given a vector of observed travel-times $\bvt{\obs} \in \re^{\sz{M}{1}}$, the SSF estimation in TTT is done by minimizing the mean squared error between the modeled (\cref{eq:sec2:bvt_product-form}) and the observed travel-times,
\begin{equation}
	\bvs\opt = \argmin_{\bvs} E(\bvs) = \norm{\bvR(\bvs) \bvs - \bvt{\obs}}^2,
	\label{eq:sec2:def_optimization_and_cost-function}
\end{equation}
where $\norm{\cdot}$ represents the traditional Euclidean ($\ell_2$) norm.

Note that this problem is non-linear: the ray paths $\bvR(\bvs)$ depends on the slowness field $\bvs$, and in turn the optimal $\bvs$ depends on the achieved ray path solution $\bvR(\bvs)$. To overcome this, the most common approach is to iteratively solve this problem. That is, given the $i$-th iteration solution $\bvs{i}$, obtain $\bvR{i} \equiv \bvR(\bvs{i})$ by solving the Eikonal equation; and given $\bvR{i}$, obtain the next iteration $\bvs{i+1}$ through
\begin{equation}
	\bvs{i+1} = \argmin_{\bvs} \norm{\bvR{i} \bvs - \bvt{\obs}}^2,
	\label{eq:sec2:bvs_i+1_basic-form}
\end{equation}
wherein $\bvR{i}$ is assumed constant for this minimization.

In this scenario, $E(\bvs)$ doesn't have a single global minimum (given $\bvR{i}$ is a wide matrix, and this is an underdetermined system). This can be addressed by adding a regularization term, yielding
\begin{equation}
	\bvs{i+1} = \argmin_{\bvs} \bar{E}(\bvs) = \norm{\bvR{i} \bvs - \bvt{\obs}}^2 + \alpha^2 \norm{\bvD \bvs}^2,
	\label{eq:sec2:bvs_i+1_regularized-form}
\end{equation}
in which $\bvD \in \re^{\sz{N}{N}}$ and $\alpha$ are a regularization matrix and parameter, respectively. The proper choice of $\bvD$ ensures a well-posed problem, improving some quality or metric in $\bvs{i+1}$ at the cost of some residual travel-time error. 

The solution to the minimization problem is given by an inverse map $\bvG{i} \in \re^{\sz{N}{M}}$. Given the quadratic formulation of \cref{eq:sec2:bvs_i+1_regularized-form}, the inverse map can be obtained by setting $\nabla \bar{E}(\bvs) = \bv0$, which leads to
\begin{equation}
	\bvG{i} = \inv*{\bvR[T]{i} \bvR{i} + \alpha^2 \bvD[T] \bvD} \bvR[T]{i}.
	\label{eq:sec2:minim_solution_literature}
\end{equation}
With this solution, $\bvs{i+1}$ can be obtained as
\begin{equation}
	\bvs{i+1} = \bvG{i} \bvt{\obs}.
    \label{eq:sec2:generic_solution_s-i+1}
\end{equation}

This formulation assumes the invertibility of $\bvR[T]{i} \bvR{i} + \alpha^2 \bvD[T] \bvD$. Given the problem's construction, $\bvR[T]{i} \bvR{i}$ can't be full-rank (since $\rank{\bvR[T]{i} \bvR{i}} = R < N$), requiring $\bvD$ to be a full-rank matrix. When this condition isn't met, non-inverting schemes have to be used, such as the popularly employed conjugate gradient method \cite{ali_opensource_2019,hormati_robust_2010,tang_travel_2024}.

As explained, the regularizing matrix $\bvD$ decides the secondary metric that is adjointed to the cost function. For example, choosing $\bvD = \bvI{N}$ (the $\sz{N}{N}$ identity matrix) results in the minimization weighting in the mean-squared value of the slowness vector \cite{phillips_traveltime_1991}, trading model accurateness for white noise rejection \cite{pan_theoretical_2015}. Alternatively, $\bvD$ can be chosen to (discretely) approximate the Laplacian of the underlying continuous slowness field \cite{ali_opensource_2019}, leading to a minimization that penalizes the SSF's curvature, resulting in a smoother field. The choice of regularization parameter $\alpha$ controls the trade-off between the primary (travel-time error) and secondary (regularization) cost functions.

\subsection{Normalized form}

Although the formulation achieved in \cref{eq:sec2:minim_solution_literature} is valid, it can be sensitive to the scale of $\bvR{i}$ (assuming fixed $\bvD$ and $\alpha$). We set $\rho = \norm{\bvR{i}}_F$ as the Frobenius norm of $\bvR{i}$ (and similar for $\delta = \norm{\bvD}_F$), and define the normalized forms $\bvR[d]{i} = \frac{\bvR{i}}{\rho}$ and $\bvD[d] = \frac{\bvD}{\delta}$.

With this modification the inverse map solution from \cref{eq:sec2:minim_solution_literature} can be written as
\begin{equation}
	\bvG{i} = \frac{1}{\rho} \inv*{\tr{\bvR[d]{i}} \bvR[d]{i} + \dot{\alpha}^2 \tr{\bvD[d]} \bvD[d]} \tr{\bvR[d]{i}},
	\label{eq:sec2:minim_solution_literature_normalized}
\end{equation}
with $\dot{\alpha} = \alpha \frac{\delta}{\rho}$. Since $\alpha$ was arbitrary, so is $\dot{\alpha}$. This allows a more robust selection of the regularization factor, for which the effects of the scale of $\bvR{i}$ are lessened. 